---
title: "Machine Learning"
author: "Joel Calixto"
date: "31 de julio de 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr);library("dplyr");library(caret);library(randomForest);
library(rattle);
```
## Machine Learning Project

##Executive Summary
From the dataset get from groupware.les.inf.puc-rio.br/har we need generate a model for the project of Machine Learning and answer 20 question from there. I generate two models by this section , one for Descision Trees and other for Random Forest.

The data was get from this url

The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.


```{r load}
library(knitr);library("dplyr");library(caret);library(randomForest);
library(rattle); library(AppliedPredictiveModeling);
library(rpart);library(rpart.plot);
library(plyr);library(dplyr);
set.seed(1307);
```

#Data load
First prepare data loading from url
```{r load1}
Url_for_train <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
Url_for_test <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
train <- "pml-training.csv"
test  <- "pml-testing.csv"
if (!file.exists("./data")) {
    dir.create("./data")
}
if (!file.exists(train)) {
    download.file(Url_for_train, destfile=paste("./data/",train), method="curl")
}
if (!file.exists(test)) {
    download.file(Url_for_test, destfile=paste("./data/",test), method="curl")
}
```

load data from local disk
```{r load2}
#reuse var for load local data in enviroment
training <- read.csv(paste("./data/","pml-training.csv"), na.strings = c("NA", "#DIV/0!", ""))
testing  <- read.csv(paste("./data/","pml-testing.csv"), na.strings = c("NA", "#DIV/0!", ""))
```

#clean data, avoid NA columns from dataSet 
```{r load3}
##str(training);str(testing)
dim(training); dim(testing)

trainning_clear<-training[, colSums(is.na(testing))==0]
testing_clear<-testing[,colSums(is.na(testing))==0]
```
remove col irrelevants for our analysis
```{r load4}
trainning_clear  <- trainning_clear[, -c(1:7)]
testing_clear   <- testing_clear[, -c(1:7)]
```
check new data set dimensions
```{r load5}
dim(trainning_clear); dim(testing_clear)
```
generate a partition for new trainning and testing dataset
```{r load6}
trainning_with_partition<-createDataPartition(trainning_clear$classe,p=0.75,list=FALSE)
trainning_for_work<-trainning_clear[trainning_with_partition,]
testing_for_work<-trainning_clear[-trainning_with_partition,]
```
#Model selection 

first probe Descision Trees
```{r load7}
#create a a model of desicion trees
DT_model<-rpart(classe~. , trainning_for_work, method="class")
#check a plot for this model
fancyRpartPlot(DT_model)
#generate a prediction - evaluate a model 
compare_dt<- predict( DT_model, testing_for_work, type="class")
#calculate the accuracy
confusion_compare_dt<-confusionMatrix(compare_dt, testing_for_work$classe)
confusion_compare_dt
```
with the descision tree model we can see an accuracy of 76.04% and a sample error of 00.31

Now we generate a Random Forest model
```{r load8}
#generate a model
RF_model<-randomForest(classe~. , trainning_for_work, method="class")
#generate a prediction - evaluate a model 
compare_rf<-predict(RF_model, testing_for_work, type = "class")
#calculate the accuracy
confusion_compare_rf<-confusionMatrix(compare_rf,testing_for_work$classe)
confusion_compare_rf
```
with the random forest we have a 99.63% of accuracy and 0.0046 

#Conclusion
For the accuracy level of Random Forest we choose this model over descisions tree

#Answers for quiz
check answers for quiz of project
Finally generate a answer for quiz
```{r load9}
final<-predict(RF_model, testing_clear)
final
```

